Extending the Advanced Forensic Format to accommodate Multiple Data Sources, Logical Evidence, Arbitrary Information and Forensic Workflow

### 2009-05-13 paper/paper.tex
simson's edit of AFF4 Ignore-this: e1cda0fe54390f6f1560d4fba53d2232

\begin{abstract}
Forensic analysis requires the acquisition and management of many
different types of evidence, including individual disk drives, RAID
sets, network packets, memory images, and extracted files. Often the
same evidence is reviewed by several different tools or examiners in
different locations. We propose a backwards-compatible redesign of the
Advanced Forensic Format---an open, extensible file format for storing
and sharing of evidence, arbitrary case related information and
analysis results among different tools. The new specification, termed
AFF4, was designed to be simple to implement, allowing the use of the
well supported Zip File format specifications for bit level file
access.
\end{abstract}

@@

\subsection{Prior Work}
In recent years there has been a steady and growing interest in the
actual file formats and containers used to store digital
evidence. Early practitioners created exact bit-for-bit copies
(commonly referred to as ``dd images''). More recently, proprietary
software systems for making and authenticating ``images'' of digital
evidence are commonly used
(e.g. \cite{safeback,ilook,encase}). PyFlag\cite{pyflag} introduced a
``seekable gzip'' format that allowed disk images to be stored in a
form that was compressed but allowed random-access to evidence data
necessary for forensic analysis. The \emph{sgzip} format does not
store any metadata at all.

@@

The Eye Witness Forensic (EWF) file format was originally developed
for Encase\cite{encase}, but then adopted by other vendors
\cite{libewf}. The EWF file format also compresses the image into 32kb
chunks which are stored back to back in groupings inside the file. The
format contains multiple tables containing relative indexes to these
compressed chunks which are also stored within the volume. EWF volumes
have a maximum size of 2Gb and therefore usually split an image across
many volumes. EWF provides for a small number of predefined metadata
fields to be stored within the file format.

Both the AFF and EWF file formats are designed to store a single
image, and any metadata implicitely refers to that image. Therefore,
these formats in themselves have a limited metadata model which is
unable to exist outside the context of the image itself.

@@

\subsection{This paper}
The AFF4 metadata model described in this work, goes beyond merely
specifying metadata as an addition to the forensic evidence
itself. Metadata forms an essential part of the file format, with the
same metadata system storing both essential information required to
read the image itself, as well as non-essential metadata. For example,
an AFF4 compliant hard disk acquisition tool should, in addition to
storing an image, store entities representing and describing the tool
itself, the operator of the tool, the user supplied parameters
supplied to the tool, and the drive being imaged.

An important advance of this work is the introduction of storage
transformation functions to the forensic storage container. Prior
works simply focused on forensically sound storage of bitstreams,
leaving the necessary activities of translating low level storage into
higher level abstractions at the aggregate block (ie. RAID), volume,
and filesystem layers in the domain of analysis tools, as transiently
constructed artifacts. In contrast AFF4 has mechanisms for describing
transformation in a flexible and concise way, allowing users to view
multiple transformations of the same data with virtually zero storage
cost. This mechanism is an important enabler for interoperable
forensic tools. For example, carved files may be described in terms of
their block allocation sequences from an image, rather than requiring
the carved file to be copied again.

This paper extends previous work on the Advanced Forensic Format (AFF)
by taking many of the concepts developed and designing a new
specification and toolset. The AFF4 format is a complete redesign of
the architecture. The new architecture is capable of storing multiple
heterogeneous data types that might arise in a modern digital
investigation, including data from multiple data storage devices, new
data types (including network packets and memory images), extracted
logical evidence, and forensic workflow. The AFF4 format extends to
become a Global Distributed Evidence Management system, while
maintaing compatibility with the old AFF file format.

We call the new system AFF4, and use the phrase AFF1 to refer to the
legacy system developed by Garfinkel et al.\footnote{Although Garfinkel
never changed the AFF bit-level specification, Garfinkel released
AFFLIB implementations with major version numbers 1, 2 and 3. We
therefore call our system AFF4 to avoid confusion.}

@@

\item AFF1 (and the EWF format) are unable to recover from loss or
  corruption. Due to the specialised nature of these formats,
  specialised recovery software must be written to recover from
  corruptions.

@@

We have also defined an information model which addresses uniquely
identifying evidence data, expressing arbitrary information and
integrating information and image data from multiple evidence volumes
at a time. This allows us to extend evidence collection and management
to a truely universal scale - allowing creation and management of a
large corpus.

@@

The AFF4 design adopts a scheme of globally unique identifiers for
identifying and referring to all evidence. We define an AFF4 specific
URN scheme, which we call the AFF4 URN. URN's of this scheme use the
namespace\cite{RFC1737} \emph{\'aff4\'} and therefore begin with the 
string(\emph{urn:aff4}). AFF4 URNs are then be made unique by use of 
a unique identifier generated
as per RFC4122\cite{RFC4122}. For example, an AFF4 URN might be
\emph{urn:aff4:bcc02ea5-eeb3-40ce-90cf-7315daf2505e}

@@

The AFF4 model treats metadata as an abstract concept which may exist
independently from the data itself. We term {\em metadata} to be a set
of statements about objects, written in triple notations (Subject,
Attribute, Value), where {\em Subject} is the URN of the object the
statement is made about. Using this system we are able to store
arbitrary attributes about any object in the AFF4
universe. Additionally, as these statements are universally scoped,
they may be stored anywhere.

@@

When the volume is loaded, the AFF4 implementation automatically loads
any properties files and populates its Universal Resolver with the
information visibile to it. AFF4 provides a mechanism to use an
external resolver stored on a MySQL database to provide for a
Universal Resolver that shares information between different instances
on the same network.

@@

\begin{itemize}
\item Most Zip implementations do not implement the Zip64
  extensions. These extensions are required to support Evidence
  Volumes larger than 2GB.

@@

This command requests an image be created directly on the evidence
server (it will be uploaded using webdav). The image is signed using
Alice's certificate and key (which might need to be unlocked). Note
that Alice does not need any hardware to obtain the image as it is
done over the network - she therefore can respond rapidly.

>>>>

\begin{abstract}
Forensic analysis requires the acquisition and management of many
different types of evidence, including individual disk drives, RAID
sets, network packets, memory images, and extracted files. Often the
same evidence is reviewed by several different tools or examiners in
different locations. We propose a backwards-compatible redesign of the
Advanced Forensic Format---an open, extensible file format for storing
and sharing of evidence, arbitrary case related information and
analysis results among different tools. The new specification, termed
AFF4, was designed to be simple to implement, bult upon the well
supported ZIP file format specification. AFF4 implements will have
downward compatability with existing AFF files.
\end{abstract}

@@

\subsection{Prior Work}
In recent years there has been a steady and growing interest in the
actual file formats and containers used to store digital
evidence. Early practitioners created exact bit-for-bit copies
(commonly referred to as ``dd images''). More recently, proprietary
software systems for making and authenticating ``images'' of digital
evidence are commonly used
(e.g. \cite{safeback,ilook,encase}). PyFlag\cite{pyflag} introduced a
``seekable gzip'' format that allowed disk images to be stored in a
form that was compressed but allowed random-access to evidence data
necessary for forensic analysis.

The Expert Witness Forensic (EWF) file format was originally developed
for Encase\cite{encase}, but then adopted by other vendors
\cite{libewf}. The EWF file format also compresses the image into 32kb
chunks which are stored back to back in groupings inside the file. The
format contains multiple tables containing relative indexes to these
compressed chunks which are also stored within the volume. EWF volumes
have a maximum size of 2Gb and therefore usually split an image across
many volumes. EWF provides for a small number of predefined metadata
fields to be stored within the file format.

@@

Both the AFF and EWF file formats are designed to store a single
image, and any metadata implicitely refers to that image. Therefore,
these formats in themselves have a limited metadata model which is
unable to exist outside the context of the image itself.
But AFF went beyond merely
specifying metadata as an addition to the forensic evidence
itself. Metadata forms an essential part of the file format, with the
same metadata system storing both essential information required to
read the image itself, as well as non-essential metadata. AFF allows
the use of arbitrary name/value pairs to store metadata, but it also
stored critical information such as sector size and device serial
number using the same metadata representation system. \verb+Aimage+,
the AFF hard  disk acquisition tool should, not only stored the image,
but also stored a description of the tool
itself, the version of AFFLIB used to create the image, the computer
on which the image was made, the operator of the tool, the user supplied parameters
supplied to the tool, all in addition to the drive being imaged.

@@

\subsection{This paper}
An important advance of this work is the introduction of storage
transformation functions to the forensic storage container. Prior
works simply focused on forensically sound storage of bitstreams,
leaving the necessary activities of translating low level storage into
higher level abstractions at the aggregate block (ie. RAID), volume,
and filesystem layers in the domain of analysis tools, as transiently
constructed artifacts. In contrast AFF4 has mechanisms for describing
transformation in a flexible and concise way, allowing users to view
multiple transformations of the same data with little additional storage
cost. This mechanism is an important enabler for interoperable
forensic tools. For example, carved files may be described in terms of
their block allocation sequences from an image, rather than requiring
the carved file to be copied again.

This paper extends previous work on the Advanced Forensic Format (AFF)
by taking many of the concepts developed and designing a new
specification and toolset. The AFF4 format is a complete redesign of
the architecture. The new architecture is capable of storing multiple
heterogeneous data types that might arise in a modern digital
investigation, including data from multiple data storage devices, new
data types (including network packets and memory images), extracted
logical evidence, and forensic workflow. The AFF4 format extends the
format to make it the basis of a global distributed evidence management system.

We call the new system AFF4, and use the phrase AFF1 to refer to the
legacy system developed by Garfinkel et al.\footnote{Although Garfinkel
never changed the AFF bit-level specification, Garfinkel released
AFFLIB implementations with major version numbers 1, 2 and 3. We
therefore call our system AFF4 to avoid confusion.} The AFF4
implementation, when it is publicly released, will be able to read
existing AFF files.

@@

@@

We have also defined an information model which addresses uniquely
identifying evidence data, expressing arbitrary information and
integrating information and image data from multiple evidence volumes
at a time. This allows us to extend evidence collection and management
to a truely universal scale---allowing creation and management of a
large corpus.

@@

The AFF4 design adopts a scheme of globally unique identifiers for
identifying and referring to all evidence. We define an AFF4 specific
URN scheme, which we call the AFF4 URN. URN's of this scheme use the
namespace\cite{RFC1737} ``\emph{aff4}'' and therefore begin with the 
string(\emph{urn:aff4}). AFF4 URNs are then be made unique by use of 
a unique identifier generated
as per RFC4122\cite{RFC4122}. For example, an AFF4 URN might be
\emph{urn:aff4:bcc02ea5-eeb3-40ce-90cf-7315daf2505e}

@@

The AFF4 model treats metadata as an abstract concept which may exist
independently from the data itself. We term {\em metadata} to be a set
of statements about objects, written in triple notations (Subject,
Attribute, Value), where {\em Subject} is the URN of the object the
statement is made about. An \emph{Attribute{ can be any kind of value, such
as the sector size of a device, a device capacity, or the name of the
person who performed an imaging operation. A \emph{Value} is the value
of the attribute. Using this system we are able to store
arbitrary attributes about any object in the AFF4
universe. Additionally, as these statements are universally scoped,
they may be stored anywhere.

@@

When the volume is loaded, the AFF4 implementation automatically loads
any properties files and populates its Universal Resolver with the
information visibile to it. AFF4 provides a mechanism to use an
external resolver as well---for example, we have implemented a
resolved that stores Attributes in a MySQL database to provide for a
Universal Resolver that shares information between different instances
on the same network.

@@

\begin{itemize}
\item The commonly available Zip implementations written in C do not
 implement the Zip64 extensions. These extensions are required to
 support Evidence Volumes larger than 2GB.

@@

This command requests an image be created directly on the evidence
server (it will be uploaded using webdav). The image is signed using
Alice's certificate and key (which might need to be unlocked). Note
that Alice does not need any hardware to obtain the image as it is
done over the network---she therefore can respond rapidly.
